

## Getting Started {#sec-get-started}

All of our examples will be based on a common data set.


Let's say you are suffering from unexplained headaches that appear somewhat randomly. You suspect they may be associated with something you eat, but you're not sure, so you've been tracking 14 weeks (98 days) worth of your own data in a spreadsheet that looks like this:

```{r, message = FALSE, warning = FALSE}
library(tidyverse, quietly=TRUE)
library(lubridate, quietly=TRUE)

```

```{r}
set.seed(1984)

x <- tibble(date=seq(from = today()-weeks(14),
                     by = "1 day", length.out = 7*14),
            headache = sample(c(TRUE,FALSE), 7*14,
                              prob = c(.05,.95),
                              replace = TRUE))

knitr::kable( head(x) ) %>% kableExtra::kable_styling()

write_csv(x,"headache-days.csv")

```

You can download a copy of this file [here](./headache-days.csv)

It's easy to add a few more variables (columns) to the dataframe: ([download](./headache-variables.csv))

```{r}

z <- function(x){
m = NULL
for(i in 1:14){
  m = c(c(rep(0,6),
              floor(runif(1,min=0,max=3))),
        m)
}

m
}
x <- tibble(date=seq(from = today()-weeks(14),
                     by = "1 day", length.out = 7*14),
            headache = sample(c(TRUE,FALSE), 7*14,
                              prob = c(.05,.95),
                              replace = TRUE),
            icecream = sample(c(TRUE,FALSE), 7*14,
                              prob = c(.10,.90),
                              replace = TRUE),
            z = runif(7*14, min = -2.5, max = .5) + 8,
            wine = z(0))

knitr::kable( head(x,10), digits = 2) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

write_csv(x,"headache-variables.csv")

```

-   `headache`: a day when I have a headache
-   `icecream`: did I eat ice cream that day?
-   `wine`: Number of glasses of wine I drank.
-   `z`: Number of hours I slept that day.



