---
title: "Personal Science Cookbook"
author: "Friends of Personal Science"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "Practical techniques to apply scientific principles to your own life."
format: html

---

# Prerequisites

## Assumptions

We'll assume some basic tools.

-   A spreadsheet like Microsoft Excel
-   The programming language R and the associated development environment RStudio

A good introduction to R is [Hands On Programming with R](https://rstudio-education.github.io/hopr/)

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
library(kableExtra)


```

## How to edit this book

This is book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. We recommend TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
