--- 
title: "Personal Science Cookbook"
author: "Richard Sprague"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "Practical techniques to apply scientific principles to your own life."
output:
  bookdown::html_document2:
    includes:
      in_header: header.html
  bookdown::gitbook:
    includes:
      in_header: header.html
---


# Prerequisites


This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

## Assumptions
We’ll assume some basic tools.

* A spreadsheet like Microsoft Excel
* The programming language R and the associated development environment RStudio

A good introduction to R is [Hands On Programming with R](https://rstudio-education.github.io/hopr/)


```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
library(kableExtra)


``` 
What is a dataframe?

Self-collected data is almost always best represented by a table of the variables you want to study and the values that you collected for each of those variables. The most common type of table is a spreadsheet, a specific form of which  in Personal Science we refer to as a data table or a _data frame_. Abbreviated “dataframe” or often just “df”, it’s a table of values and variables that always has the same form:

* columns are variables: the parameters you want to study
* rows are observations: each incident of data you collected.

It’s important to get in the habit of this row/column approach to data collection because, as you’ll see, all of our tools assume that data will come in a dataframe format. 


Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
