---
title: "Machine learning"
---

When you are tracking lots of variables, it can be difficult to spot exactly which ones are driving an effect. 

In this example, we've fudged the data so that headaches always happen only on days with less than 6 hours of sleep.

Also, headache days always have high stress. But sometimes we have high stress even without a headache.


```{r}
#| warning: false
#| message: false
library(tidyverse)
library(lubridate)

set.seed(1984)
WEEKS <- 140

z <- function(x){
m = NULL
for(i in 1:WEEKS){
  m = c(c(rep(0,6),
              floor(runif(1,min=0,max=3))),
        m)
}

m
}
x <- tibble(date=seq(from = today()-weeks(WEEKS),
                     by = "1 day", length.out = 7*WEEKS),
            headache = sample(c(TRUE,FALSE), 7*WEEKS,
                              prob = c(.05,.95),
                              replace = TRUE),
            stress = sample(c("low","medium","high"),
                            size = 7*WEEKS,
                            replace = TRUE),        
            icecream = sample(c(TRUE,FALSE), 7*WEEKS,
                              prob = c(.10,.90),
                              replace = TRUE),
            z = runif(7*WEEKS, min = -2.5, max = .5) + 8,
            wine = z(0))

x$headache <- x$z<6  # make headaches on days with low sleep
x[x$headache, "stress"] <- "high"  # make stress on headache days

knitr::kable( head(x,10), digits = 2) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

Now make a linear model that shows `r nrow(x)` rows

```{r}

m <- lm(headache~icecream+z+wine+stress, data = x)
summary.lm(m)
p <- predict.lm(m)

plot(p)

```

Those triple dots ("***") in the right-hand column for Coefficients indicate items of very high significance.

In this case, as expected, there seems to be a strong relationship between both sleep and stress.