---
title: "Machine learning"
---

When you are tracking lots of variables, it can be difficult to spot exactly which ones are driving an effect. 

In this example, we've fudged the data so that headaches always happen only on days with less than 6 hours of sleep.

Also, headache days always have high stress. But sometimes we have high stress even without a headache.


```{r}
#| warning: false
#| message: false
library(tidyverse)
library(lubridate)

set.seed(1984)
WEEKS <- 140

z <- function(x){
m = NULL
for(i in 1:WEEKS){
  m = c(c(rep(0,6),
              floor(runif(1,min=0,max=3))),
        m)
}

m
}
tracking_table <- tibble(date=seq(from = today()-weeks(WEEKS),
                     by = "1 day", length.out = 7*WEEKS),
            headache = sample(c(TRUE,FALSE), 7*WEEKS,
                              prob = c(.05,.95),
                              replace = TRUE),
            stress = sample(c("low","medium","high"),
                            size = 7*WEEKS,
                            replace = TRUE),        
            icecream = sample(c(TRUE,FALSE), 7*WEEKS,
                              prob = c(.10,.90),
                              replace = TRUE),
            z = runif(7*WEEKS, min = -2.5, max = .5) + 8,
            wine = z(0))

tracking_table$headache <- tracking_table$z<6  # make headaches on days with low sleep
# tracking_table[tracking_table$stress=="high"] <- "high"  # make stress on headache days

knitr::kable( head(tracking_table,10), digits = 2) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

Now make a linear model using the `r nrow(tracking_table)` rows from our dataframe.

```{r}

m <- lm(headache~icecream+z+wine+stress, data = tracking_table)
summary(m)
```

Those triple dots ("***") in the right-hand column for Coefficients indicate items of very high significance.

In this case, as expected, there seems to be a strong relationship between both sleep and stress.

A quick yet dramatic way to visualize this uses the `stat_smooth` function of `ggplot`:


```{r}
tracking_table %>% ggplot(aes(y=headache, x=z)) + geom_point() + stat_smooth(formula = y~x, method = "lm", color = "red") +
facet_grid(headache ~ stress)

```
We clearly see that headaches always occur on days when Z<6.

And here's what happens if we change the data so headaches always happen when stress is high.

```{r}

tracking_table[tracking_table$stress == "high",]$headache <- TRUE

tracking_table %>% ggplot(aes(y=headache, x=z)) + geom_point() + stat_smooth(formula = y~x, method = "lm", color = "red") +
facet_grid(headache ~ . + stress)


```

As expected, the plot shows headaches either when `z<6` or when stress is high.

